Results_paper1

### 2019-04-12 Paper1_results.Rmd
Data_push

# Libraries

@@

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(heatwaveR)
```

@@

* See the Dectect_MCS.Rmd script for more info. The datasets loaded here are only for temperatures collected between the years 1992-2016. However if a longer time series is required changes may  be made to the Dectect_MCS.Rmd script (line 136) . MCS using a long term time series to detect any extreme events occuring within the region

@@

```{r}
filtered_years <- function(df){
  upwelling<- df %>% 
  filter(year(date) %in% seq(2002, 2015)) %>% 
  drop_na()
}

SACTN_fyears <- filtered_years(df = SACTN_overlap) 
MUR_fyears <- filtered_years(df = insitu_MUR) 
OISST_fyears <- filtered_years(df = insitu_OISST) 
G1SST_fyears <- filtered_years(df = insitu_G1SST) 
K10_fyears <- filtered_years(df = insitu_K10)

# save(SACTN_fyears, file = "Data/temp_2014_2015/SACTN_fyears.RData")
# save(MUR_fyears, file = "Data/temp_2014_2015/MUR_fyears.RData")
# save(OISST_fyears, file = "Data/temp_2014_2015/OISST_fyears.RData")
# save(G1SST_fyears, file = "Data/temp_2014_2015/G1SST_fyears.RData")
# save(K10_fyears, file = "Data/temp_2014_2015/K10_fyears.RData")
```

@@

```{r}
# When doing a wavelet analyses on the SACTN dataset the "insitu_temp" column needed to be renamed to "temp" for the rest of the code to run. The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)
# SACTN_fyears <- SACTN_fyears %>%
#   dplyr::rename(temp = insitu_temp)

@@

for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures/new",plot_name,"_wavelets.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun)
```

>>>>

# Loading libraries

@@

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(forecast)
library(astrochron)
library(WaveletComp)
library(data.table)
library(heatwaveR)
library(viridis)
library(ggrepel)
library(plyr)
library(maptools)
library(sp)
library(geosphere)
library(marmap)
library(PBSmapping)
library(scales)
library(grid)
library(gridExtra)
library(fossil)
library(mapproj)
library(rgeos)

## Functions
source("functions/earthdist.R")
```

@@

@@

```{r}
filtered_years <- function(df){
  upwelling<- df %>% 
  filter(year(date) %in% seq(2010, 2014)) %>% 
  drop_na()
}

SACTN_fyears <- filtered_years(df = SACTN_overlap) 
MUR_fyears <- filtered_years(df = insitu_MUR) 
OISST_fyears <- filtered_years(df = insitu_OISST) 
G1SST_fyears <- filtered_years(df = insitu_G1SST) 
K10_fyears <- filtered_years(df = insitu_K10) 
```

@@

```{r}
# When doing a wavelet analyses on the SACTN dataset the "insitu_temp" column needed to be renamed to "temp" for the rest of the code to run. The rest of this code is kept constant for each of the datasets (Remotely-sensed SST and In-situ seawater temperature)

SACTN_fyears <- SACTN_fyears %>%
  dplyr::rename(temp = insitu_temp)

@@

for (i in 1:length(PN_wave)) {
  attributes(PN_wave[[i]]) <- c(attributes(PN_wave[[i]]), ref = names(PN_wave)[i])
}
# x <- PN_wave[[3]]
plot_fun <- function(x, plot_name = attributes(x)$ref) {
  png(filename = paste0("Figures",plot_name,"_wavelets.png"),
      width = 800, height = 600, units = "px", pointsize = 12, bg = "white")
  wt.image(x, siglvl = 0.05, col.contour = "black", color.key = "quantile",
           timelab = "Days", verbose = FALSE, useRaster = TRUE,
           periodlab = "Period", lwd = 1, graphics.reset = FALSE,
           main = plot_name)
  dev.off()
}
ldply(PN_wave, plot_fun) 
```

############################################################################################################################################

# Calculating a shore transect
The aim of this is to find the temperature collected at different distances from the coast (distances = 15km, 25km, 35km and 50km). Then create wavelet analyses (or another analyses) where I compare how the intensity and duration of upwelling vary at different distances from the coastline. So would I have a dataset with temperature at 15km then at 25km then at 35km and then at 50km and so on...?

```{r}
load("Data_P1/site_list_sub.Rdata")
west <- site_list_sub
west$coast <- "west" # Chnages wc to west

load("Data_P1/africa_coast.RData")

## Downloading the bathy data from NOAA
# Download mid-res bathymetry data
# sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
# sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
# colnames(sa_bathy) <- c("lon", "lat", "depth")
# sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
# save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")

# Loading in the newly downloaded bathymetry data
load("Data_P1/bathy/sa_bathy.RData")

# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
shore.normal.transect <- function(site, width = 2){
  # Find the site on the coastline and it's nearest neighbour points
  coords <- data.frame(lon = site$lon, lat = site$lat)
  coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
  coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
  coords3 <- coords3[2:1,1:3]
  # Define the shore normal transect bearing
  heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
  if(heading >= 360){
    heading <- heading-360
  } else {
    heading <- heading
  }
  heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
  return(heading2)
}

# Creating the transects
site_transects <- data.frame()
for(i in 1:length(west$site)){
 site <- west[i,]
 site_transect <- shore.normal.transect(site, 2)
 site_transects <- rbind(site_transects, site_transect)
}

# Manually correcting Sea Point and Kommetjie
site_transects$heading[5:6] <- 290 
save(site_transects, file = "Data_P1/site_transects.RData")
load("Data_P1/site_transects.RData")

# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
# It then extracts a lat/ lon point every X kilometres until reaching a specified isobath

transect.pixel.isobath <- function(site, distance = 25000, isobath = 999){ # Only at a distance of 25000km
  # Extract coordinates
  coords <- data.frame(lon = site$lon, lat = site$lat)
  # Find lon/ lats every X metres until the chosen isobath is reached
  pixels <- data.frame()
  deep <- 999
  while(deep > isobath){
    # Find depth at distance step
    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distance))
    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
    bathy2 <- sa_bathy[sitesIdx,]
    bathy2 <- bathy2[complete.cases(bathy2$depth),]
    bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
                         heading = site$heading, depth = bathy2$depth)
    pixels <- rbind(pixels, bathy3)
    coords <- coords2
    deep <- bathy2$depth
  }
  pixels2 <- pixels[pixels$depth > -999,] # This may remove the only row of data
  if(nrow(pixels2) < 1){
    pixels2 <- data.frame(site, depth = NA)
  }else{
    pixels2 <- pixels2
  }
  return(pixels2)
}

# Pixel points
site_pixels <- data.frame()
for(i in 1:length(west$site)){
  site <- site_transects[i,]
  site_pixel <- transect.pixel.isobath(site, 25000, -900)
  site_pixels <- rbind(site_pixels, site_pixel)
}

save(site_pixels, file = "Data_P1/site_pixels.RData")
load("Data_P1/site_pixels.RData")

# Bounding box
  # Only one is made in order to know how large the the geom_point() squares should be made to match
test <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
                   xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
                   ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
                   ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
```

## Create a visualisation of the sites at the different distances from the coastline
****Fix: Fill to show the temperature variation at Benguela coast| also fix the site names*

load("Data_P1/south_africa_coast.RData")
load("Data_P1/bathy/sa_bathy.RData")
names(south_africa_coast)[1] <- "lon"

# Manually divide up coastline
wc <- south_africa_coast[291:410,]

# Define plotting parameters
sa_lats <- c(-37, -27); sa_lons <- c(14, 34)

# Plotting with pixel boxes
plot1 <- ggplot() +
 geom_contour(data = bathy[bathy$depth >= -250,], aes(x = lon, y = lat, z = depth),
               colour = "grey20", alpha = 0.7, size = 0.2, binwidth = 200, na.rm = TRUE, show.legend = FALSE) +
  stat_contour(data = sa_bathy[sa_bathy$depth < -250,], aes(x = lon, y = lat, z = depth, alpha = ..level..),
               colour = "grey20", size = 0.1, binwidth = 1000, na.rm = TRUE, show.legend = FALSE) +
  geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group), 
               size = 0.5, colour = "black", fill = "white") +
   # geom_text(data = west, aes(x = lon, y = lat, label = site),
   #          size = 1.8, colour = "black") +
  # geom_polygon(data = wcBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
  #              colour = "bisque1", fill = "bisque1") +
  geom_point(data = west, aes(x = lon, y = lat), alpha = 0.8, size = 1.6) +
  geom_text(data = west, aes(x = lon, y = lat, label = site), 
                     size = 6, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, 
                     segment.alpha = 0.4, force = 0.1) +
  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "black", shape = 0, alpha = 0.8, size = 2.1) +
  scale_alpha_continuous(breaks = c(-200, -1000, -2000, -3000, -4000, -5000),
                         guide_legend(title = "depth (m)")) +
  scale_fill_gradient(low = "dodgerblue4", high = "dodgerblue", na.value = "steelblue4", 
                      breaks = c(-1000, -2000, -3000, -4000, -5000),
                      guide_legend(title = "depth (m)")) +
   annotate("text", label = "INDIAN\nOCEAN", x = 34.00, y = -35.0,
             size = 6.0, angle = 0, colour = "black") +
    annotate("text", label = "ATLANTIC\nOCEAN", x = 14.00, y = -35.0,
             size = 6.0, angle = 0, colour = "black") +
    annotate("text", label = "Benguela", x = 15.0, y = -31.7,
             size = 4.0, angle = 302, colour = "black") +
    annotate("text", label = "Agulhas", x = 32.0, y = -31.7,
             size = 4.0, angle = 50, colour = "black") +
  labs(title = NULL, x = NULL, y = NULL) +
  theme(legend.key = element_rect(colour = NA, size = 0.2),
        legend.key.height = unit(0.4, "cm"),
        legend.background = element_blank()) +
  coord_cartesian(xlim = sa_lons, ylim = sa_lats) 
plot1
```

# Upwelling correlated with wind patterns
## Loading in the Wind data

```{r}
wind_1 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind1/wind1.txt", na.strings="") 
colnames(wind_1) <- c("station_number", "station_name", "date", "hour", "sub", "speed", "dir")

wind_2 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind2/wind2.txt", na.strings="")
colnames(wind_2) <- c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir")

wind_3 <- read.delim("Data_P1/Wind_data/wind_data.txt(SAWS)/wind3/wind3.txt", na.strings="")
colnames(wind_3) <- c("station_number", "station_name", "date", "hour", "sub" ,"speed", "dir")

# Slecting the important columns for each of the datasets
wind_fix <- function(df){
wind <- df%>% 
  select(station_name, date, hour, dir, speed)
}
wind_1 <- wind_fix(df = wind_1)
wind_2 <- wind_fix(df = wind_2)
wind_3 <- wind_fix(df = wind_3)

## Renaming the sites within the wind datasets to match the name of the sites at which seawater temperature was collected
## The wind data was obtained from the SAWS and the wind stations used were the closes stations to which temperature was collected

renaming_sites_1 <- function(df) {
  sites <- df %>%
    mutate(temp_sites = ifelse(station_name %in% c("CAPE COLUMBINE"), "St Helena Bay",        
                           ifelse(station_name %in% c("KOINGNAAS"), "Hondeklipbaai",
                                ifelse(station_name %in% c("PORT NOLLOTH"), "Port Nolloth",
                                       ifelse(station_name %in% c("LAMBERTSBAAI NORTIER"), "Lamberts Bay",
                                              ifelse(station_name %in% c("LANGEBAANWEG AWS"), "Saldanha Bay
","Error"))))))
  return(sites)
}

wind_sitesmatched_1 <-  renaming_sites_1(df = wind_1)
wind_sitesmatched_1 <- wind_sitesmatched_1[-c(121352, 121353, 379892, 379893, 609324, 609325, 843506, 843507, 1014585), ]

renaming_sites_2 <- function(df) {
  sites <- df %>%
    mutate(temp_sites = ifelse(station_name %in% c("ROBBENEILAND"), "Koeberg Basin",        
                           ifelse(station_name %in% c("GEELBEK"), "Yzerfontein",
                                ifelse(station_name %in% c("DASSEN ISLAND"), "Dassen Island",
                                       ifelse(station_name %in% c("ATLANTIS"), "Koeberg Basin","Error")))))
  return(sites)
}

wind_sitesmatched_2 <-  renaming_sites_2(df = wind_2)
wind_sitesmatched_2 <- wind_sitesmatched_2[-c(228372, 228373, 313441, 313442, 364881, 364882, 557392), ]

renaming_sites_3 <- function(df) {
  sites <- df %>%
    mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Kommetjie",  
                           ifelse(station_name %in% c("CAPE TOWN - ROYAL YACHT CLUB"), "Sea Point",
                           ifelse(station_name %in% c("CAPE TOWN TABLE BAY"), "Sea Point","Error"))))
  return(sites)
}

wind_sitesmatched_3 <-  renaming_sites_3(df = wind_3)
wind_sitesmatched_3 <- wind_sitesmatched_3[-c(119317, 119318, 196551, 196552, 346759), ]

## CAPE TOWN SLANGKOP may be used for Kommetjie and for Houtbay
wind_3_HoutBay <- wind_3 %>% 
  filter(station_name == "CAPE TOWN SLANGKOP") %>% 
  mutate(temp_sites = ifelse(station_name %in% c("CAPE TOWN SLANGKOP"), "Hout Bay","Error"))

wind_data <- rbind(wind_3_HoutBay,wind_sitesmatched_3,wind_sitesmatched_2,wind_sitesmatched_1)
wind_data <- wind_data %>% 
  na.omit()
# save(wind_data, file = "Data_P1/wind_data.RData")

# The wind data

wind_data <- wind_data %>% 
  as.numeric(dir)

wind_data <- wind_data %>%
  mutate(date = as.Date(date)) %>%
  group_by(temp_sites, date) %>%
  summarise(mean_dir = mean(as.numeric(dir),
            mean_speed = mean(as.numeric(speed))))

save(wind_data, file = "Data/wind_daily.RData")

# load("Data/wind_daily.RData") # Loading the daily wind data for each of the sites
```

## Upwelling indeces

```

In particular, trends will be observed to be different at ocean and coastal locations and we hypothesize that
those differences are due to changes in upwelling intensity.

- Detemining the upwelling indeces in R
- How does upwelling indeces vary at the different distances along the coast
- Positive indeces display a positive upwelling.
